### 2014-11-05 D1-3.tex
Extended Threat Scenarios

Live+Gov Privacy Dashboard gives citizens back the Control

@@

\chapter{Summary}\label{chap:Summary}

>>>>

@@

\chapter{Summary}\label{chap:Summary}
* Discussed Legal and ethical aspects of privacy
* Made Privacy Security Analysis of Live+Gov System
* Presented several Privacy Protection Measures that have been implemented.
* Most importantly the Privacy Dashboard, which gives the citizen control over their data.

### 2014-11-05 Implementations.tex
Extended Threat Scenarios

\textbf{What Personal Information do we Collect?}\\
The Mobile Sensor Collection Application collects the following data from your mobile phone:\\
- Sensor Data from Accelerometer (50Hz)\\
- Sensor Data from GPS sensors (~every 30 sec)\\
- Daily Human Activities (walking, standing, sitting, etc.) (every 30sec)\\
- Used Service Lines (bus, train, tram) (every 2 minutes)\\

All collected data is transferred to our data center.

The Live+Gov Services does NOT store or collect names and email addresses.

@@

\textbf{When is the data deleted?}\\
The full data-set will be stored until January 2015.

Afterwards we will delete Names and Email addresses and anonymize
the sensor data and processing results.  The anonymized data will
be used by the the Consortium Partners for research purposes.

\textbf{For which purposes is the data collected?}\\
The collected data will be used for personalization of user experience of the mobile application.
Moreover, we will analyize travel patterns in the Helsinki Urban region for general research and for improvement and optimization of the
infrastructure provided by HSL.

\textbf{Identity of Data Controller}\\
The data is controlled by Heinrich Hartmann <hartmann@uni-koblenz.de>

The following organizations have access to the collected data:
\begin{enumerate}
\item University of Koblenz-Landau
\url{http://www.uni-koblenz-landau.de/}\\
Universitätsstraße 1\\
56070 Koblenz

\item Mattersoft
\url{http://www.mattersoft.fi/}
Hämeenkatu 13, 33100 Tampere, Finland\\
+358 10 3225000

\item Centre for Research and Technology Hellas (CERTH)
\url{http://mklab.iti.gr/contact}\\
6th km Charilaou-Thermi Road\\
P.O. Box 60361,\\
57001 Thermi-Thessaloniki, Greece
\end{enumerate}

@@

There you have the possibility to:\\
- View all recordings made\\
- View the raw data collected for each recording\\
- Download the data for each recording\\
- Delete records

@@

@@

@@

In the case that such information should be linked to the user-id
stored in the trip table we provide a python script that randomizes
the user-ids and thereby removes all links that might have existed
before.

@@

The Gaussian Noise anonymization is a basic method is usually mixed
well with other methods.

>>>>

\textbf{What Personal Information do we Collect?}\\
The Mobile Sensor Collection Application collects the following data from your mobile phone:\\
- Sensor Data from Accelerometer\\
- Sensor Data from GPS sensors\\
- Daily Human Activities (walking, standing, sitting, etc.)\\
- Used Service Lines (bus, train, tram)

All collected data is transferred to our data center. The Live+Gov
Services does NOT store or collect names and email addresses.

@@

@@

There you have the possibility to:\\
- View all recordings made\\
- Delete individual recordings\\
- View the raw data collected for each recording\\
- Download the data for each recording\\
- View the results of the applied data processing

\textbf{When is the data deleted?}\\
The full data-set will be stored until January 2016.

Anonymized artifacts of this data are stored longer than that and will
be used by the Consortium Partners for research purposes.

\textbf{For which purposes is the data collected?}\\
The collected data will be used for personalization of user experience
of the mobile application.  Moreover, we will analyze travel patterns
in the Helsinki Urban region for general research and for improvement
and optimization of the infrastructure provided by HSL.

@@

\textbf{Who has access to the data?}
The following organizations inside the Live+Gov Consortium have access
to the collected data:
\begin{enumerate}
\item University of Koblenz-Landau
\url{http://www.uni-koblenz-landau.de/}\\
Universitätsstraße 1\\
56070 Koblenz

\item Mattersoft \\
\url{http://www.mattersoft.fi/} \\
Hämeenkatu 13, 33100 Tampere, Finland\\
+358 10 3225000

\item Centre for Research and Technology Hellas (CERTH)\\
\url{http://mklab.iti.gr/contact}\\
6th km Charilaou-Thermi Road\\
P.O. Box 60361,\\
57001 Thermi-Thessaloniki, Greece
\end{enumerate}

@@

\textbf{Identity of Data Controller}\\
The data is controlled by Dr. H.  Hartmann <hartmann@uni-koblenz.de>.

@@

In the case that such information should be linked to the user-id
stored in the trip table we provide a python script (cf. Figure
\ref{fig:privacypy}) that randomizes the user-ids and thereby removes
all links that might have existed before. The scripts allows as a
parameter a general SQL query that selects a sub-set of the trips that
shall be anonymized.

@@

The Gaussian Noise anonymization, even with a low noise radius of 5m,
disables the possibility to correlate GPS data with data from cameras
to identify a single person. It is typically used in conjunction with
other method to provide a maximum of anonymity.

### 2014-11-05 PrivacyProtection-2014-09-09.tex
finished threat eval table
Extended Threat Scenarios

\section{Privacy Definition and Taxonomy}
\label{sec:taxonomy}

@@

Please refer to Section \ref{sec:taxonomy} for more details.

@@

\textbf{Unhappy Employees.} Susceptible for social engineering.
%TODO: Add description

@@

% Extend to general missing privacy wareness of citizen, system provider and local authority
\textbf{Unaware Monitoring Subjects.}
We define privacy as one's ability to control information about oneself.
In order to do that, monitored subjects need to know, that they are monitored, who monitors them, what information is recorded and for what purposes.
Subjects who are not aware of these things cannot effectively preserve control and thus lose their privacy.

%TODO
- No Privacy Policy
- No Information about Processing, Access rules, 3rd party disclosure

@@

%3. Surveillance
%---------------
%Description:   External infiltrates infrastructure in order to spy on citizen
%COI:           Privacy vs. External Power
%Vulnerability: Insecure Infrastructure, Insecure Communication
%Affected Data:         Data transferred and stored in data center.
%
%Likeliness: Medium (Government*), Low (Third parties)
%Loss: High (Control is lost, potential harm by due to Government action)
%Risk:      6/4
%
%* Government, has also other tools at it's disposal
%
%
\textbf{Surveillance}
Lorem Ipsum

%4. Information Leak
%-------------------
%Description: Like an external person in Scenario (2) an employee of the service provider has selfish interests to gain money, make political statements or harm his employer.
%In order to pursue this interest he can steal personal data and sell it or release it to the public.
%
%COI:         Privacy vs. Financial Profit
%             Reputation of Service Provider vs. Political Activism
%             Reputation of Local Authority vs. Political Activism
%Vuln.:       Unhappy Employee, Lax Access Rules.
%Affected Data:       Data stored in data center.
%
%Likeliness: Low
%Loss: High (public leak)
%Risk: 4
%
\textbf{Information Leak}
Lorem Ipsum

%5. Social Engineering
%---------------------
%Description: An External can manipulate Local Authorities or Service Providers to give them personal information that he can then sell or disclose.
%
%COI:         Privacy vs. Financial Profit
%             Reputation of Service Provider vs. Political Activism
%             Reputation of Local Authority vs. Political Activism
%Vuln.:       Unhappy Employee, Lax Access Rules.
%Affected Data:       Data stored in data center.
%
%Likeliness: Low
%Loss: High
%Risk: 4
%
\textbf{Social Engineering}
Lorem Ipsum

@@

In this section we will associate to every identified threat a corresponding risk.
Recall from \ref{sec:GrimmModel} that a risk is the expected loss that is associated to the threat.
Therefore, we have to quantify the likeliness of the threat to occure and the harm or loss done in this case.
The quantification of likeliness will be solely based on rough judgment of the authors.
The quantification of loss, will be made in a two step process.
For each threat listed in table \ref{firgure:Threats}, we have analyzed the affected personal data of the citizen.
For each possible data type (e.g. GPS) we analyze the impact on the seven different types of privacy in Section \ref{sec:SensorPrivacyImpact}.
In combination we can quantify roughly the impact of each threat on the citizens privacy.

Both evaluations are necessarily fraught with a high level of uncertainty.

Loss Calibration
\begin{itemize}
\item 3: High. Leak of information to Peers which impacts citizen.
  E.g. spouse gets GPS data and finds out cheating.
  Subjects is added to No-Fly list due to Government spying.
  Public leak of private information.
\item 2: Medium. Undisclosed processing of personal data or disclosure to third parties that are unrelated to subject.
\item 1: Low. Loss of control over data.
\item 0: None
\end{itemize}

Likeliness is measured on a the following scale
\begin{itemize}
\item 0 (impossible)
\item 1: maybe occure once in 1 million cases.
\item 2: maybe occure once in 100
\item 3: maybe occure once in 10
\item 4: always the case.
\end{itemize}
The risk scale is defined as addition of the loss and likeliness scales.\footnote{
Note, that loss and likeliness scales have a logarithmic character, so that that addition of those scales corresponds to multiplication of the usual scales.
}

@@

\subsubsection{Security Requirement Specification}
%TODO SKIP?
TODO: Formulate requirements that address the high risk scenarios.

%\subsection{Step 3. Plan Development}

% \subsection{Legal Privacy Requirements} \ref{sec:LEGAL_REQ}

%\subsubsection{Security Measures}

% Seven Cs are now in separate Document

>>>>

\section{Definition and Types of Privacy}
\label{sec:taxonomy}

@@

We refer to Section \ref{sec:taxonomy} for more details.

@@

\textbf{Unhappy Employees.}

An Employee that is frustrated with his situation for a long time
period constitutes a security vulnerability. On the one hand he might
want to harm his employer directly, on the other he is increasingly
susceptible for social engineering.

@@

%TODO: Extend to general missing privacy wareness of citizen, system provider and local authority
\textbf{Unaware Monitoring Subjects.}
We define privacy as one's ability to control information about
oneself.  In order to do that, monitored subjects need to know, that
they are monitored, who monitors them, what information is recorded
and for what purposes.  Subjects who are not aware of these things
cannot effectively preserve control and thus lose their privacy.  This
vulnerability expresses itself as the lack of information material
like a Privacy Policy including concise information about applied
processing steps, access rules and disclosure to 3rd parties.

@@

%3. Surveillance
%---------------
%COI:           Privacy vs. External Power
%Vulnerability: Insecure Infrastructure, Insecure Communication
%Affected Data:         Data transferred and stored in data center.
%
%Likeliness: Medium (Government*), Low (Third parties)
%Loss: High (Control is lost, potential harm by due to Government action)
%Risk:      6/4
%
%* Government, has also other tools at it's disposal
%
%
\textbf{Surveillance} External infiltrates infrastructure in order to
obtain direct about the citizen. In this scenario the external is
supposed to have some direct relationship to the citizen which
motivates his interest to obtain personal information. Examples could
be a public institution that wants to gain information about planned
activities of the citizens (e.g. Nixon's Watergate scandal or the
recent prosecution of Guardian journalists by GHCQ). Another example
is an insurance company that seeks to get information about the
citizens life-style in relation to the insured risk, like car
accidents or health hazards.

In this threat the privacy interest of the citizen is in conflict with
the aspirations for power over the citizen by the externals.

%4. Information Leak
%-------------------
%COI:         Privacy vs. Financial Profit
%             Reputation of Service Provider vs. Political Activism
%             Reputation of Local Authority vs. Political Activism
%Vuln.:       Unhappy Employee, Lax Access Rules.
%Affected Data:       Data stored in data center.
%
%Likeliness: Low
%Loss: High (public leak)
%Risk: 4
%

\textbf{Information Leak}
Like an external person in Scenario (2) an employee of the service provider has selfish interests to gain money, make political statements or harm his employer.
In order to pursue this interest he can steal personal data and sell it or release it to the public.

%5. Social Engineering
%---------------------
%Description: 
%COI:         Privacy vs. Financial Profit
%             Reputation of Service Provider vs. Political Activism
%             Reputation of Local Authority vs. Political Activism
%Vuln.:       Unhappy Employee, Lax Access Rules.
%Affected Data:       Data stored in data center.
%
%Likeliness: Low
%Loss: High
%Risk: 4
%
\textbf{Social Engineering}
An External can manipulate Local Authorities or Service Providers to
give them personal information that he can then sell or disclose.

@@

In this section we will associate to every identified threat a corresponding risk.
Recall from \ref{sec:GrimmModel} that a risk is the expected loss that is associated to the threat.
Therefore, we have to quantify the likeliness of the threat to occure and the harm or loss done in this case.
The quantification of likeliness will be solely based on rough judgment of the authors.
The quantification of loss, will be made in a two step process.
For each threat listed in table \ref{firgure:Threats}, we have analyzed the affected personal data of the citizen.
For each possible data type (e.g. GPS) we analyze the impact on the seven different types of privacy in Section \ref{sec:SensorPrivacyImpact}.
In combination we can quantify roughly the impact of each threat on the citizens privacy. Both evaluations are necessarily fraught with a high level of uncertainty.

For the quantification of the loss in case of a threat scenario we use the following rough calibration:
\begin{itemize}
\item 3: High. Leak of information to Peers which impacts citizen.
  E.g. spouse gets GPS data and finds out cheating.
  Subjects is added to No-Fly list due to Government spying.
  Public leak of private information.
\item 2: Medium. Undisclosed processing of personal data or disclosure to third parties that are unrelated to subject.
\item 1: Low. Loss of control over data.
\item 0: None
\end{itemize}

For the quantification of likeliness the following scale is used:
\begin{itemize}
\item 4: always the case.
\item 3: maybe occure once in 10
\item 2: maybe occure once in 100
\item 1: maybe occure once in 1 million cases.
\item 0 (impossible)
\end{itemize}

The quantification of the risk, we add the values for loss and
likeliness of the corresponding threats. Note, that loss and
likeliness scales have a logarithmic character, so that that addition
of those scales corresponds to multiplication of the usual scales.

@@

\subsubsection{Security Recommendations and Requirements}
\textbf{Legal Requirements}

Privacy Policy

Processing Journal
\begin{itemize}
\item name of the controller
\item purpose of processing
\item description of the data categories
\item recipients of the data if disclosed
\item transfer to third countries
\item description of security of processing
\end{itemize}
The journal is available for anybody to access via Email.
Reporting to Data Protection Authority (Bundesamt für Datenschutz).
